Comparing Distance Profiles Using Violin Plots and Shift Function

### 2019-04-22 Scripts/comparing_distance_profiles_using_violin_and_shift_plots.Rmd
editted script

violin.gm12878.10kb.smc <- ggplot(annotdata.smc, aes(x=ordered(y, levels=rev(unique(y))), y=Smc3ab9263_dist, color=y)) + 
  geom_violin(trim=FALSE)+
  geom_boxplot(width=0.1, fill="white")+
  coord_flip() +
  scale_color_discrete(name="Class of\nGenomic Bin",
                         labels=c("Did not overlap\nw/ flanked TAD boundary", "Overlapped w/\nflanked TAD boundary"))+
  xlab("Class of Genomic Bin") +
  ylab("Log2 Distance") +
  ggtitle("SMC3") +
  #annotate("text", x=7, y=.2, label= paste("P-Value:", perm.pval, sep = " "), size=5) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 15),
    axis.text.y = element_blank(),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    legend.text=element_text(size=15),
    legend.title=element_text(size=20),
    plot.title = element_text(size=20,
                              hjust=.5))

@@

shift.gm12878.10kb.smc <- ggplot(out, aes_string(x=xplot, y="difference")) +
  geom_hline(yintercept=0,linetype=2,alpha=0.5) + # x=0 reference line
  geom_vline(xintercept=xintercept,linetype=2,alpha=0.5) + # y=median reference line
  geom_errorbar(aes(ymin=ci_lower, ymax=ci_upper), colour="black", width=.15) +
  geom_line(colour="black", linetype="solid", size=1.5) +
  geom_point(colour="black", size=4, shape=21, fill="white") + #999999
  xlab("Quantiles for bins that\nOverlapped w/flanked TAD boundaries") +
  ylab("Quantile differences between bin class") +
  theme_bw() +
  theme(axis.text.x = element_text(size=14),
        axis.text.y = element_text(size=14),
        axis.title.x = element_text(size=16,face="bold"),
        axis.title.y = element_text(size=16,face="bold")) +
  scale_y_continuous(limits = ylim)

ggarrange(violin.gm12878.10kb.smc,
          shift.gm12878.10kb.smc,
          ncol=1,
          nrow=2,
          common.legend = TRUE,
          legend = "top")
```

>>>>

violin.gm12878.10kb.smc <- ggplot(annotdata.smc, aes(x=ordered(y, levels=rev(unique(y))), y=Smc3ab9263_dist, color=y)) + 
  geom_violin(trim=FALSE)+
  geom_boxplot(width=0.1, fill="white")+
  coord_flip() +
  scale_color_discrete(name="Class of\nGenomic Bin",
                         labels=c("Did not overlap\nw/ flanked TAD boundary", "Overlapped w/\nflanked TAD boundary"))+
  xlab("Class of genomic bin") +
  ylab("Log2 distance") +
  ggtitle("SMC3") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 15),
    axis.text.y = element_blank(),
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    legend.text=element_text(size=15),
    legend.title=element_text(size=15),
    legend.position="top",
    plot.title = element_text(size=20,
                              hjust=.5))

@@

shift.gm12878.10kb.smc <- ggplot(out, aes_string(x=xplot, y="difference")) +
  geom_hline(yintercept=0,linetype=2,alpha=0.5) + # x=0 reference line
  geom_vline(xintercept=xintercept,linetype=2,alpha=0.5) + # y=median reference line
  geom_errorbar(aes(ymin=ci_lower, ymax=ci_upper), colour="black", width=.10) +
  geom_line(colour="black", linetype="solid", size=1.5) +
  geom_point(colour="black", size=4, shape=21, fill="white") + #999999
  xlab("Quantiles for bins that Overlapped\nw/flanked TAD boundaries") +
  ylab("Quantile differences") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15),
        axis.text.y = element_text(size=15),
        axis.title.x = element_text(size=15),
        axis.title.y = element_text(size=15)) +
  scale_y_continuous(limits = ylim)

####ZNF143

```{r}
annotname <- gsub(paste(c("`", 
             "Gm12878", 
             "-", 
             #"_", 
             #"dist", "count", "perc",
             "Haib", 
             "Sydh",
             "Uta",
             "Uw",
             "Broad",
             "Uchicago"), collapse = "|"), "", "Gm12878-Znf143166181ap-Sydh_dist")

annotdata.znf143 <- train_smote[, c(1, which(names(train_smote)==annotname))]

#ks.test(annotdata.znf143[which(annotdata.znf143$y=="No"),2], annotdata.znf143[which(annotdata.znf143$y=="Yes"),2], alternative = "two.sided")

perm.pval <- pvalue(oneway_test(annotdata.znf143[,2] ~ annotdata.znf143[,1],
                     data=annotdata.znf143,
                     alternative="greater", 
                     distribution=approximate(B=10000)))[1]

plottitle <- paste(gsub(paste(c("Gm12878",
                                  "-","_",
                                  "dist","perc","count",
                                  "Haib",
                                  "Sydh",
                                  "Uta",
                                  "Uw",
                                  "Broad",
                                  "Uchicago"), 
                                collapse = "|"), 
                          "",
                          "Gm12878-Znf143166181ap-Sydh_dist"))

violin.gm12878.10kb.znf143 <- ggplot(annotdata.znf143, aes(x=ordered(y, levels=rev(unique(y))), y=Znf143166181ap_dist, color=y)) + 
  geom_violin(trim=FALSE)+
  geom_boxplot(width=0.1, fill="white")+
  coord_flip() +
  scale_color_discrete(name="Class of\nGenomic Bin",
                         labels=c("Did not overlap\nw/ flanked TAD boundary", "Overlapped w/\nflanked TAD boundary"))+
  xlab("Class of genomic bin") +
  ylab("Log2 distance") +
  ggtitle("ZNF143") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 15),
    axis.text.y = element_blank(),
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    legend.text=element_text(size=15),
    legend.title=element_text(size=15),
    legend.position="top",
    plot.title = element_text(size=20,
                              hjust=.5))

out <- shifthd(annotdata.znf143$Znf143166181ap_dist[which(annotdata.znf143$y=="No")],
               annotdata.znf143$Znf143166181ap_dist[which(annotdata.znf143$y=="Yes")],
               nboot=200)

shift.gm12878.10kb.znf143 <- ggplot(out, aes_string(x=xplot, y="difference")) +
  geom_hline(yintercept=0,linetype=2,alpha=0.5) + # x=0 reference line
  geom_vline(xintercept=xintercept,linetype=2,alpha=0.5) + # y=median reference line
  geom_errorbar(aes(ymin=ci_lower, ymax=ci_upper), colour="black", width=.10) +
  geom_line(colour="black", linetype="solid", size=1.5) +
  geom_point(colour="black", size=4, shape=21, fill="white") + #999999
  xlab("Quantiles for bins that Overlapped\nw/flanked TAD boundaries") +
  ylab("Quantile differences") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15),
        axis.text.y = element_text(size=15),
        axis.title.x = element_text(size=15),
        axis.title.y = element_text(size=15)) +
  scale_y_continuous(limits = ylim)

####CTCF

```{r}
annotname <- gsub(paste(c("`", 
             "Gm12878", 
             "-", 
             #"_", 
             #"dist", "count", "perc",
             "Haib", 
             "Sydh",
             "Uta",
             "Uw",
             "Broad",
             "Uchicago"), collapse = "|"), "", "Gm12878-Ctcf-Broad_dist")

annotdata.ctcf <- train_smote[, c(1, which(names(train_smote)==annotname))]

#ks.test(annotdata.ctcf[which(annotdata.ctcf$y=="No"),2], annotdata.ctcf[which(annotdata.ctcf$y=="Yes"),2], alternative = "two.sided")

perm.pval <- pvalue(oneway_test(annotdata.ctcf[,2] ~ annotdata.ctcf[,1],
                     data=annotdata.ctcf,
                     alternative="greater", 
                     distribution=approximate(B=10000)))[1]

plottitle <- paste(gsub(paste(c("Gm12878",
                                  "-","_",
                                  "dist","perc","count",
                                  "Haib",
                                  "Sydh",
                                  "Uta",
                                  "Uw",
                                  "Broad",
                                  "Uchicago"), 
                                collapse = "|"), 
                          "",
                          "Gm12878-Ctcf-Broad_dist"))

violin.gm12878.10kb.ctcf <- ggplot(annotdata.ctcf, aes(x=ordered(y, levels=rev(unique(y))), y=Ctcf_dist, color=y)) + 
  geom_violin(trim=FALSE)+
  geom_boxplot(width=0.1, fill="white")+
  coord_flip() +
  scale_color_discrete(name="Class of\nGenomic Bin",
                         labels=c("Did not overlap\nw/ flanked TAD boundary", "Overlapped w/\nflanked TAD boundary"))+
  xlab("Class of genomic bin") +
  ylab("Log2 distance") +
  ggtitle("CTCF") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 15),
    axis.text.y = element_blank(),
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    legend.text=element_text(size=15),
    legend.title=element_text(size=15),
    legend.position="top",
    plot.title = element_text(size=20,
                              hjust=.5))

out <- shifthd(annotdata.ctcf$Ctcf_dist[which(annotdata.ctcf$y=="No")],
               annotdata.ctcf$Ctcf_dist[which(annotdata.ctcf$y=="Yes")],
               nboot=200)

shift.gm12878.10kb.ctcf <- ggplot(out, aes_string(x=xplot, y="difference")) +
  geom_hline(yintercept=0,linetype=2,alpha=0.5) + # x=0 reference line
  geom_vline(xintercept=xintercept,linetype=2,alpha=0.5) + # y=median reference line
  geom_errorbar(aes(ymin=ci_lower, ymax=ci_upper), colour="black", width=.10) +
  geom_line(colour="black", linetype="solid", size=1.5) +
  geom_point(colour="black", size=4, shape=21, fill="white") + #999999
  xlab("Quantiles for bins that Overlapped\nw/flanked TAD boundaries") +
  ylab("Quantile differences") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15),
        axis.text.y = element_text(size=15),
        axis.title.x = element_text(size=15),
        axis.title.y = element_text(size=15)) +
  scale_y_continuous(limits = ylim)

####RAD21

```{r}
annotname <- gsub(paste(c("`", 
             "Gm12878", 
             "-", 
             #"_", 
             #"dist", "count", "perc",
             "Haib", 
             "Sydh",
             "Uta",
             "Uw",
             "Broad",
             "Uchicago"), collapse = "|"), "", "Gm12878-Rad21-Haib_dist")

annotdata.rad21 <- train_smote[, c(1, which(names(train_smote)==annotname))]

#ks.test(annotdata.rad21[which(annotdata.rad21$y=="No"),2], annotdata.rad21[which(annotdata.rad21$y=="Yes"),2], alternative = "two.sided")

perm.pval <- pvalue(oneway_test(annotdata.rad21[,2] ~ annotdata.rad21[,1],
                     data=annotdata.rad21,
                     alternative="greater", 
                     distribution=approximate(B=10000)))[1]

plottitle <- paste(gsub(paste(c("Gm12878",
                                  "-","_",
                                  "dist","perc","count",
                                  "Haib",
                                  "Sydh",
                                  "Uta",
                                  "Uw",
                                  "Broad",
                                  "Uchicago"), 
                                collapse = "|"), 
                          "",
                          "Gm12878-Rad21-Haib_dist"))

violin.gm12878.10kb.rad21 <- ggplot(annotdata.rad21, aes(x=ordered(y, levels=rev(unique(y))), y=Rad21_dist, color=y)) + 
  geom_violin(trim=FALSE)+
  geom_boxplot(width=0.1, fill="white")+
  coord_flip() +
  scale_color_discrete(name="Class of\nGenomic Bin",
                         labels=c("Did not overlap\nw/ flanked TAD boundary", "Overlapped w/\nflanked TAD boundary"))+
  xlab("Class of genomic bin") +
  ylab("Log2 distance") +
  ggtitle("RAD21") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 15),
    axis.text.y = element_blank(),
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    legend.text=element_text(size=15),
    legend.title=element_text(size=15),
    legend.position="top",
    plot.title = element_text(size=20,
                              hjust=.5))

out <- shifthd(annotdata.rad21$Rad21_dist[which(annotdata.rad21$y=="No")],
               annotdata.rad21$Rad21_dist[which(annotdata.rad21$y=="Yes")],
               nboot=200)

shift.gm12878.10kb.rad21 <- ggplot(out, aes_string(x=xplot, y="difference")) +
  geom_hline(yintercept=0,linetype=2,alpha=0.5) + # x=0 reference line
  geom_vline(xintercept=xintercept,linetype=2,alpha=0.5) + # y=median reference line
  geom_errorbar(aes(ymin=ci_lower, ymax=ci_upper), colour="black", width=.10) +
  geom_line(colour="black", linetype="solid", size=1.5) +
  geom_point(colour="black", size=4, shape=21, fill="white") + #999999
  xlab("Quantiles for bins that Overlapped\nw/flanked TAD boundaries") +
  ylab("Quantile differences") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15),
        axis.text.y = element_text(size=15),
        axis.title.x = element_text(size=15),
        axis.title.y = element_text(size=15)) +
  scale_y_continuous(limits = ylim)

###Plots

```{r}
ggarrange(violin.gm12878.10kb.smc,violin.gm12878.10kb.znf143,violin.gm12878.10kb.ctcf,violin.gm12878.10kb.rad21,
          shift.gm12878.10kb.smc,shift.gm12878.10kb.znf143,shift.gm12878.10kb.ctcf,shift.gm12878.10kb.rad21,
          ncol=4,
          nrow=2,
          common.legend = TRUE,
          legend = "top",
          labels = c("A","B","C","D","","","",""),
          font.label = list(size=20))
